---
output: github_document
# bibliography: reference.bib  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
    fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# R package UQforecasts

## Overview

This package accompanies the paper "Uncertainty Quantification in Forecast Comparisons" by Marc-Oliver Pohle, Tanja Zahn and Sebastian Lerch.

The goal of the package is to provide *valid simultaneous inference in forecast comparisons* in a multi-dimensional setup while preventing multi-comparison problems. It can be used to estimate measures such as skill scores, relative accuracy, the expected scores themselves and the difference in expected scores. The latter is the primary object of interest in the popular Diebold-Mariano tests  of equal predictive accuracy. In addition, the package will generate confidence bands of level $1 - \alpha$ that are simultaneous over all dimensions, e.g., over forecasting methods, forecast horizons, variables and/or locations (in the case of meteorological forecasts).

The main function of the package is called `conf_bands`. This function estimates the measure of interest and generates simultaneous confidence bands of level $1-\alpha$. Since the output is a nested list, we also provide a function called `list_to_df` that transforms the output to a data frame.

For convenience, we also added some customary plotting functions:

* `plot_horizontal`: Plot the measure and the confidence bands horizontally, e.g., across forecast horizons, and potentially facet by another variable.
* `plot_vertical`: Plot the measure and the confidence bands vertically, e.g., across different variables when there is only one forecast horizon.
<!-- * `plot_map`: Create a map, e.g., across different locations. -->
<!-- * `plot_map_seq`: Create a map with a sequential color scheme. -->

## Installation


To install the package from Github and load it, run the following `R` code:

```{r, results = FALSE,  message=FALSE, eval = FALSE}
install.packages("devtools")
library(devtools)
install_github("TanjaZahn/UQforecasts")
library(UQforecasts)
```



## Example 1

### Creating toy data

Before we introduce the main function of the package, we will create some toy data that is also attached to this package. We create some artificial scores, e.g., the squared errors or the CRPS, of two forecasting methods ($M = 2$) for an evaluation sample of size $N = 100$. We consider $H = 4$ forecast horizons and $V = 3$ variables. For this toy example, we just draw the scores for each $t = 1,..,N$ as absolute values from a normal distribution. We set the mean of the scores for the first method equal to $v +h$, where $v = 1,...,3$ and $h=1,...,4$. Thus, the method has the highest mean, i.e. performs worst on average, for the third variable and the fourth forecast horizon. For the second method, we set the mean equal to $2+h$, that is, the mean of the scores still increase with the forecast horizon. Compared to method 1, method 2 performs better for variable 1 and worse for variable 3, on average. The performance for variable 2 should be similar.

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Load additional packages to create toy data
library(tidyverse)

# Set seed for replication
set.seed(42)

# Parameters
N <- 100 # size of the evaluation sample
H <- 4 # maximum forecast horizons
V <- 3 # number of variables
M <- 2 # number of methods

# Make a data frame
df_toy <- expand.grid(t = 1:N, h = 1:H, v = 1:V)

# Create some artificial scores like, e.g., the squared error, for each time period.
df_toy <- df_toy %>% 
  group_by(h, v) %>% 
  mutate(score1 = abs(rnorm(N, v + h, 1)),
         score2 = abs(rnorm(N, 2 + h, 1))) %>% 
  ungroup()

# For convenience, we will transform this into the long format:
df_toy <- df_toy %>% pivot_longer(starts_with("score"),
                          names_to = "m", names_prefix = "score",
                          values_to = "score") %>% 
  mutate(m = as.integer(m))
df_toy

```


### Transforming data into a nested list

In order to apply the main function `conf_bands`, we need to transform our data set containing the scores into a nested list using `lapply()`. First, we need to iterate over methods (even if $M = 1$), then over time periods. The order of the remaining dimensions (horizons and variables in this example) does not matter. The last element of the list structure has to be a numeric value, that is, the score for method $m$, at time $t$, forecast horizon $h$ and for variable $v$.

For later convenience, we will iterate over named vectors because the names allow us to convert the results back to a data frame easily: However, the names are not necessary for `conf_bands`. 

```{r}

# Create named vectors for iteration.
methods <- setNames(1:M, c("m1", "m2")) # forecasting methods
t_eval <- setNames(1:N, paste0("t", 1:N)) # periods in evaluation sample
vars <- setNames(1:V, paste0("v", 1:V)) # variables
horizons <- setNames(1:H, paste0("h", 1:H)) # forecast horizon

# Create nested list of scores
scores <- 
  lapply(methods, function(mm) 
    lapply(t_eval, function(tt) 
      lapply(horizons, function(hh) 
        lapply(vars, function(vv){
          
          df_toy %>% filter(m == mm, t == tt, h == hh, v == vv) %>% pull(score)
  
}))))

# Example: Score for method 2, in period 50, at forecast horizon 2, for variable 3
scores[[2]][[50]][[2]][[3]]

```

## The main function: `conf_bands`

The main function `conf_bands` allows to compute four different metrics with confidence bands: `"ES"` (expected scores), `"SS"` (skill scores), `"RA"` (relative accuracy), and `"D"` (difference in expected scores). As an example, we will estimate the skill scores and the expected scores themselves for the toy data. 

We estimate the skill score of method 1 relative to method 2 (`m_bench = length(scores) = 2`) and construct Bonferroni confidence bands of level $1-\alpha = 0.95$ that are simultaneous with respect to the variables and the forecast horizons. The output of `conf_bands` has the same structure as the input list except that the time dimension is missing. Further, when calculating skill scores, one method is used as benchmark, which results in `length(skill) = M - 1 `.

```{r}
# Estimate the skill with confidence bands
skill <- conf_bands(scores, metric = "SS" , m_bench = length(scores), type = "bonferroni", 
                    B = 1000, l = 3*floor(length(scores[[1]])^(1/4)), alpha = 0.05)

# Example: Skill score with conf. bands of method 1 (relative to benchmark), at horizon 2, for variable 3
skill[[1]][[2]][[3]]

```

If the input list is named, `conf_bands`will preserve the names. Then, `list_to_df` can be used to transform the results back into a data frame. We specify the new column names in the same order as in the nested list `skill`: `dimnames = c("m", "v", "h")` or, e.g., `dimnames = c("method", "var", "horizon")`. Optionally, we can also choose to get rid of the prefixes in the the columns specified in `dims_numeric`.

```{r}
df_skill <- list_to_df(skill, dimnames = c("m", "h", "v"),  dims_numeric = c("h", "v"))
df_skill

```

As another example, we estimate the expected scores for both methods.
```{r}
expscore <- conf_bands(scores, metric = "ES", type = "bonferroni", 
                    B = 1000, l = 3*floor(length(scores[[1]])^(1/4)), alpha = 0.05)

df_expscore <- list_to_df(expscore, dimnames = c("m", "h", "v"),  dims_numeric = c("h", "v"))
df_expscore

```



### Plotting results

For convenience, we included some custom functions for plotting the results. Often, we are interested in comparing forecasts across several forecast horizons, that is, we want to plot them horizontally using the function `plot_horizontal`. First, we plot the skill score (`y = "SS`) across horizons (`xvar = "h"`). In this case, our bands are not only simultaneous with respect to the horizons but also across variables. Thus, we additionally facet by `"v"`:

```{r}
plot_horizontal(df_skill, y = "SS", xvar = "h", facet_by = "v", theme = theme_bw())

```

As expected, the skill score is positive for variable 1, meaning that method 1 outperforms method 2 for the first variable by approximately 32.4\% at $h = 1$. With increasing forecast horizon, the difference becomes less relevant but the lower bound of the confidece bands is above zero at 7.7\%. That is, at all considered forcast horizons, method 1 has signifcant gains of considerable magnitude as compared to method 2. For variable 2, both methods perform comparable without the trajectory being significanlty different from zero. For variable 3, method 2 performs better than method 1, with the gains becoming less relevant with increasing forecast horizon.

When plotting the expected scores themselves (`y = "ES"`), we additionally use `color_by` for the methods. Moreover, we get rid of the vertical line marking zero by using `markzero = FALSE`.

```{r}
plot_horizontal(df_expscore, y = "ES",  xvar = "h", facet_by = "v", color_by = "m", markzero = FALSE, theme = theme_bw())

```

## Example 2

In this section, we will use the data we have created in the previous section and aggregate scores over forecast horizons. This example is similar to the previous section except that we will illustrate plotting the results via `plot_vertical`.

### Aggregating scores and estimating expected scores

```{r}
# Aggregate
df_agg <- df_toy %>% group_by(t, v, m) %>% mutate(score_agg = sum(score)) %>% filter(row_number() == 1) %>% ungroup()

# Make nested list
scores_agg <- lapply(methods, function(mm) lapply(t_eval, function(tt) lapply(vars, function(vv){
  
  df_agg %>% filter(m == mm, t == tt, v == vv) %>% pull(score_agg)
  
})))

# Estimate expected scores
expscore_agg <- conf_bands(scores_agg, metric = "ES", type = "bonferroni", 
                    B = 1000, l = 3*floor(length(scores[[1]])^(1/4)), alpha = 0.05)

# Make data frame
df_expscore_agg <- list_to_df(expscore_agg, dimnames = c("m", "v"),  dims_numeric = c("v"))


```

### Plot results

Now, we plot the expected scores (aggregated over horizons) with vertical confidence bands for the three variables (`xvar = "v"`) and the two methods (`color_by = "m"`):

```{r}
plot_vertical(df_expscore_agg, xvar = "v", y = "ES", color_by = "m", markzero = FALSE, , theme = theme_bw())

```

